---
title: "Assignment1"
output:
  pdf_document: default
  html_document: default
date: "2023-02-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Answering Questions
### Part 1

3. Discuss your approach and whether the dot or bar approach will be more effective in communicating the information required to support your article.  (about half page )

Ans: I believe that the dot plot approach will be more effective in understanding the required concept i.e to compare the recent CPI percentage of our country along with 8 other countries and one organization. Bar chart is not that effective as we are not required to show the amount or ratio of Location variable. Instead our only concern is the final percentage which dot plots represent better than bar charts as bar charts focuses on the quantity which is not a requirement. 

### Part2



## Importing and Reading the Data
```{r}
library(dplyr)
library(ggplot2)

data <- read.csv('D:/Anagha/Studies/Assignment/DV/1/Counsumer_Price_Index.csv')

```


```{r}

```

## Filtering the Data
```{r}

filtered_data <- data %>% filter(Time == "2022-12")
filtered_data <- filtered_data %>% filter(Subject %in% c('Total', 'Food', 'Total_Minus_Food_Energy'))

#unique(filtered_data$Location)

filtered_data <- filtered_data %>% filter(Location %in% c('AUT', 'BEL', 'CAN', 'CZE', 'DNK', 'FIN', 'FRA', 'DEU', 'IRL' ,'EU27_2020'))

IRL <- filtered_data%>% filter(Location %in% "IRL")
IRL
```

## Implement the dot plot 
```{r}
dot_plot <- ggplot(filtered_data, aes(x = reorder(Location, -Percentage), y = Percentage)) + 
  geom_point(size = 3, aes(shape=Subject,color=Location%in%"IRL"))+
  #scale_shape_manual(values = c(16,16,16,16,16,16,15,16,16,17)) +
  scale_color_manual(values = c("#808080", "#FF0000")) +
  theme_minimal()+
  ylab("Percentage CPI")+
  xlab("Location")+
  labs(title="Recent Percentage CPI of Ireland with 8 other countries and 1 organization")+
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
        plot.margin=margin(t=3, r=15, b=3, l=3, unit="pt"),
        panel.grid.major.x=element_line(colour="grey60", linetype="dotted", size=0.07),
        axis.text.x = element_text(angle = 30, vjust = 0.5),
        plot.title = element_text( size = 10, face = "bold"),
        legend.position = "bottom",
        legend.box.background = element_rect(colour="#808080"))+
  guides(color="none")
  
  

dot_plot
```


```{r}

```

## Implement the bar plot 
```{r}
bar_plot <- ggplot(filtered_data, aes(x = reorder(Location,-Percentage), y = Percentage, fill = Subject, alpha = Location%in%"IRL")) + 
  geom_bar(stat = "identity", position = "dodge") +
  scale_alpha_manual(values=c(0.5, 1))+
  theme_minimal()+
  scale_y_continuous(
    breaks = c(0, 5, 10, 15, 20, 25, 30),
    expand = c(0, 0),
    name = "Percentage CPI") +
  xlab("Location")+
  labs(title="Recent Percentage CPI of Ireland with 8 other countries and 1 organization")+
  theme(axis.title.y = element_text(margin = margin(r = 20)),
        axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
        plot.margin=margin(t=3, r=15, b=3, l=3, unit="pt"),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = "bottom",
        legend.key.size = unit(0.3, 'cm'), #change legend key size
        legend.key.height = unit(0.3, 'cm'), #change legend key height
        legend.key.width = unit(0.3, 'cm'), #change legend key width
        legend.title = element_text(size=11), #change legend title font size
        legend.text = element_text(size=8),
        plot.title = element_text( size = 10, face = "bold"),
        legend.box.background = element_rect(colour="#808080"))+
  guides(alpha="none")

bar_plot
```

3. Comment upon changes you have made since part 1

Ans: In the dot plot, initially the x-axis was not ordered but now it is ordered. I changed the position of legend to bottom. Colors were used to show different subject while shapes were used to differentiate Ireland, other 8 countries, and one group of organization. The color for the Ireland was set to red and rest of the locations were set to dark grey that made Ireland easy to find.

Similarly, the x-axis labels of bar plots were also ordered and using alpha, the location of Ireland was made brighter than the rest locations. Rest everything is kept the same.

```{r}

```


```{r}

```


```{r}

```

